\subsection{Incremental Entity Extraction}
\todo{maybe join with clustering}
\todo{incremental kbp}

As anticipated in \Cref{sec:incompleteness}, \acp{kr} and \acp{kb}\todo{ensure also kr are incomplete in sec:incompleteness} suffer from incompletess, and one of the use of \ac{ie} is to populate them (\acf{kbp})~\cite{entityorientedsearch2018balog}, for instance, with new movies extracted from newspapers.

The entities missing from a \ac{kr} are usually referred to as \emph{NIL entities}~\cite{entityorientedsearch2018balog}, where \emph{NIL} may stand for ``not in lexicon''~\cite{ILIEVSKI2020100565}. Another interpretation is that it simply refers to the English word ``nil'' that means ``nothing''~\cite{cambridge_nil}.

\begin{definition}[NIL Prediction]
    Given an entity mention $m$ and a reference \ac{kr}, \emph{NIL prediction} is the subtask of \acl{nel} that identifies whether $m$ referes to an entity that does not exist in the \ac{kr}, i.e., a \emph{NIL entity}~\cite{entity_linking_with_kb_2015}. 
\end{definition}
Through this work we will also speak of \emph{NIL mentions}, referring to mentions of NIL entities. 

This task has also been called with different names: ``unlinkable mention prediction''~\cite{entity_linking_with_kb_2015}, ``out-of-knowledge-Base mention discovery''~\cite{dong_reveal_the_unknown_2023}, ``out-of-knowledge-graph entity detection''~\cite{moller2024entity}, ``unknwon entity discovery''~\cite{kassner-etal-2022-edin}, and ``emerging entity discovery''~\cite{hoffart_discovering_emergin_entities_2014}.

\textcite{zhou-etal-2024-gendecider} distinguishes between NIL prediction and ``none of the candidates'' detection, arguing that \emph{NIL} means the referred entity is not in the \ac{kr}, while ``none of the candidates'' means the referred entity has not been retrieved without assuming it is not present in the \ac{kr}. However, in practise, NIL prediction is often performed by predicting whether the top-ranked entity, returned by the \ac{nel} ranking step, is correct or not~\cite{entity_linking_with_kb_2015}. It is impractical and prohibitive to compare a mention with every entity in a \ac{kr} for ensuring the mentions is NIL. Thus, in this work, we consider the preactical assumption that if a mention does not refer to the top-ranked candidate(s), returned by the \ac{nel} system, the mention is NIL. Under this assumption there is no difference between NIL prediction and none-of-the-candidates prediction.

\todo{definition of NIL clustering}
For addressing the incompleteness problem of \acp{kr}, once identified, NIL mentions that refer to the same NIL entity can be clustered to gather more contextual information, useful for representing the new entity. Indeed, \textcite{kassner-etal-2022-edin} achieved better results when constructing the entity representation from clusters of NIL mentions with respect to just using individual NIL mentions.

\begin{definition}[NIL Clustering]
    \emph{NIL clustering} is the task that aims at clustering NIL mentions that refer to the same NIL entity~\cite{ji2011tac}.
\end{definition}

NIL prediction has frequently been overlooked in \ac{nel}
systems, as many approaches exclude NIL mentions from their evaluation datasets and
eventually leave this task to future work~\cite{blink}. In fact, among the 38 approaches\todo{verify the published version}
compared by the survey \textcite{Sevgili2022NeuralEL} only 8 addressed NIL prediction. \todo{update for clustering too}

% dico che le TAC hanno dato importanza al NIL
Early work on NIL prediction dates back to 2006, with \textcite{bunescu-pasca-2006-using} using a threshold-based approach.
Later, the Text Analysis Conference (TAC)\footnote{\url{https://tac.nist.gov/}} fostered the research in NIL prediction by introducing NIL prediction in its \acf{kbp} track starting from 2009 \cite{mcnamee2009overview}. Starting from 2011, participants were also required to cluster together those NIL mentions that referred to the same entity \cite{ji2011tac}. Organized by NIST\footnote{\url{https://www.nist.gov/}}, TAC aimed to advance research in \ac{nlp} and related areas by offering large-scale test collections and standardized evaluation procedures. Since 2008, it has been held annually, inspiring a variety of studies on NIL prediction.

Also, from the 2015 edition, NIL prediction and NIL clustering have been included in the Named Entity rEcognition and Linking (NEEL) challenge~\cite{rizzo2017lessons}, a challenge focusing on microposts, i.e., posts from social media platforms such as x.com\footnote{\url{https://x.com/}}.


%% nil pred

\ac{nel} with NIL prediction can be framed as a classification problem with a reject option, which corresponds to NIL. According to \textcite{Sevgili2022NeuralEL}, four main strategies are typically used:
\begin{enumerate}
\item Assigning NIL whenever candidate generation produces no possible entities.
\item Applying a threshold to the linking score, below which the mention is treated as unlinkable.
\item Introducing NIL as an additional candidate during the ranking stage, effectively creating a new class in the classification problem.
\item Training a separate binary classifier on mention-entity pairs, often with extra features such as the top linking score or \ac{ner} labels, to decide whether a mention is linkable or not.
\end{enumerate}

Some studies \cite{Zhang2012TheNE, Rao2013EntityLF} argue that methods \ref{nilm-threshold} and \ref{nilm-add-class} are essentially equivalent, although the latter avoids manual threshold tuning.

%% nil approaches
Several approaches to NIL prediction have been proposed across the years.
Joint approaches to \ac{ner}, \ac{nel}, and NIL prediction have been proposed. \textcite{martins-etal-2019-joint} used an LSTM-based model to perform NER, NEL, and NIL prediction jointly. Earlier, \textcite{Fahrni2012HITSMA,Fahrni2013hits} jointly performed NEL, NIL prediction, and NIL clustering.

Threshold-based methods are also common. Systems in TAC 2017, such as those by \cite{BernierColborne2017CRIMsSF,Jiang2017SRCBED}, follow this approach, as do earlier works \cite{tan2015buptteam,Byrne2013UCDIA,huang2013msr,dalton2013umass}. These systems typically apply a score threshold to decide when a mention should be assigned NIL.

Another strategy is to add NIL as an additional class~\cite{Broscheit_2019,Rao2013EntityLF,Li2017AHM,klang2019overview,chen2013casia,barrena2013ubc,zhang2012nlprir,mcnamee2011cross}. For example, \textcite{Broscheit_2019} relied on BERT with a classification layer.

Binary classification approaches have also been widely studied~\cite{moreno_combining_word_entity_embeddings_2017,Yang2017TheTS,yang2013buptteam,Pink2013SYDNEYCA,zhou2013description,monahan2012lorify}. Some methods leveraged additional statistical or score-based features. \textcite{Yang2017TheTS}, for instance, incorporated the second-best candidate’s score and the standard deviation of the top-$k$ scores.

Finally, some systems rely on alternative methods. For example, \textcite{Paikens-EtAl:2017:TAC-KBP} used estimated coherence scores between the linked entities in the document.
\todo{add methods from moller}

% Advancements in NIL prediction continued also in parallel to the TAC conference~\cite{gottipati-jiang-2011-linking,hoffart_discovering_emergin_entities_2014,hoffart_awakens_2016,wu2016exploring}, even if \textcite{gottipati-jiang-2011-linking} experimentation was run on the TAC-KBP 2010 dataset~\cite{ji2010overview}. 

%%

More recently, since 2022, the interest in NIL prediction has motivated multiple studies to improve the popular dual-encoder (bi-encoder and cross-encoder) \ac{nel} approach~\cite{blink} for NIL prediction~\cite{ijckg2022,kassner-etal-2022-edin,aydin-etal-2022-find,Heist2023nastylinker,zhu-etal-2023-learn}.

\textcite{ijckg2022}---that is one of the contributions of my Ph.D., and is better detailed in \Cref{ijckg2022}---proposed a benchmark dataset, an evaluation procedure, and a baseline pipeline for \emph{incremental \acl{nel}}. In this scenario, given $N$ batches of textual documents $b_0..b_N$, corresponding to $N$ time-steps, and an initial \ac{kr} $KR_0$, each time a batch of documents $b_k$ is processed, the identified NIL entities are added to the \ac{kr} $KR_{k+1}$, which will be used as the refernce \ac{kr} when processing the batch $b_{k+1}$.

Similarly, \textcite{kassner-etal-2022-edin} propose a dataset derived from Wikipedia and newspapers data, divided into two parts. The first part contains the documents exising before a time $t$, while the second the documents created after. The authors also proposed a pipeline-based system for end-to-end incremental \acl{nel}, based on the bi-encoder. They use a threshold for NIL prediction and cluster all the identified mentions (including the non-NIL ones) together with the entities from the \ac{kr}, based on embeddings similarities and thresholds. The clusters without entities from the \ac{kr} corresponds to NIL entities.

Other two datasets have been proposed for incremental \ac{nel} in 2022~\cite{tempel_2022,nilk_dataset_2022}.
\textcite{tempel_2022} proposal is composed of ten yearly snapshots of Wikipedia from 2013 to 2022. NILK~\cite{nilk_dataset_2022} instead considers Wikidata as the reference \ac{kb}, using English Wikipedia for the and is composed of two snapshots: Wikidata 2017 and 2022.

% main entity linking approach also include nil 
Still in 2022, \textcite{ayoola-etal-2022-refined} proposed ReFinED, ``an efficient zero-shot-capable approach to end-to-end \ac{nel}'', based on the bi-encoder architecture, and additionally able to perform NIL prediction. However, NIL prediction was not studied in the evaluation.
\textcite{aydin-etal-2022-find} studied \ac{el} with NIL prediction in the funding domain, to better aid funding organization in tracking the outcome of their expenses.\todo{avoid repetitions with nel}

\textcite{zhu-etal-2023-learn} proposed both a dataset and an improvement of the dual encoder \ac{nel} architecture, which performs NIL prediction applying a threshold on the mention-entity similarity. The dataset is created starting from the aliases of ambiguous seed entities and taking Wikipedia extracts mentioning the aliases balancing between when the mention is a hyperlink and when not. This because the author want to distinguish between NIL entities and ``non-entity phrases'', i.e., common phrases usually not considered entities, such as ``the householder''~\cite{zhu-etal-2023-learn}. About 10,000 entries were collected and manually annotated.

Also \textcite{Heist2023nastylinker} proposed a NIL-prediction aware \ac{nel} method, based on the dual encoder architecture~\cite{blink}, that uses a clustering phase, followed by a conflict resolution phase, to identify NIL entities as clusters without entities from the \ac{kr}. They evaluate on the NILK~\cite{nilk_dataset_2022} dataset.

\textcite{agarwal-etal-2022-entity} method involves building a graph whose nodes correspond to mention and entities. The edges are weighted according to the mention-mention or mention-entity similarity, calculated using bi-encoder embeddings. Next, edges are pruned with heuristics that maintain the validity of contraints, e.g., a cluster contains at most one entity.

\textcite{dong_reveal_the_unknown_2023} propose another extension of the dual encoder architecture of \textcite{blink} for NIL prediction. The authors add a NIL entity to the \ac{kr} described by ``It is a NIL option'' and use it with both bi-encoder and cross-encoder---including it in top-k entity retrieved, if not present, before the cross-encoder step. They add an additional linear layer to the cross-encoder with a sigmoid function to compute a NIL probability score and jointly train the cross-encoder for \ac{nel} and NIL prediction. Finally, they experiment different strategies for building NIL-aware datasets, namely manual labeling, pruning---which implies removing a portion of entities, labeling their mentions to NIL---or \ac{kr} versioning---based on snapshots of the same \ac{kr}, taken at different times.

\todo{improve this review with some categorization,... not just a mere drop of summaries}

\textcite{moller2024entity} addressed both NIL prediction and clustering focusing on Wikidata and combining the bi-encoder architecure with knowledge graph embeddings (TransE~\cite{transe_2013}). They retrieve a set of 100 candidates with TF-IDF and fuzzy search, then they rank them according to a) the mention-entity similarity calculated with bi-encoder embeddings of the mention in context and of the Wikidata entity description, b) popularity of the candidate entity, c) the average similarity with past linkings calculated with TransE entity embeddings. In a second stage, they cluster---with DBSCAN~\cite{dbscan_1996}---the NIL mentions based on the similarities from the mention encoder embeddings and the TransE embeddings combined with a trained linear layer.

An alternative approach is suggested by \textcite{Gu_Qu_Wang_Huai_Yuan_Gui_2021} which focuses on mircoposts and avoids the dual-encoder paradigm. The authors modeled the ranking phase of \ac{nel} with NIL prediction task as multi-choice \acl{qa}, giving the mention with candidate entities with description and the NIL option as input to a BERT model.

% llm based
With respect to \ac{llm}-based approaches~\cite{zhou-etal-2024-gendecider,ding-etal-2024-chatel}, they perform NIL prediction in the candidate ranking step of \ac{nel} by mapping the task to a multi-choice question answering problem\todo{extend this categorization to el and nil pred}, including the NIL option, in the form of ``None of the Candidates'' and ``None of the entity match'', respectively.


\todo{trattare insieme nil pred e clustering}






