%\pagestyle{mystyle}
\chapter{Introduction}

\section{Introduction}

% While benchmark data\dots

% Knowledge-intensive domains. what are they?

% information need? which ones?

% extraction and access

Knowledge-intensive domains are fields that demand specialized expertise, deep understanding, and extensive information processing, such as law, healthcare, finance, and scientific research~\cite{niu-etal-2025-tree,cambridge-knowledge-intensive}. In these areas, decisions rely heavily on the expert's domain knowledge~\cite{kg-intensive-nlu-xai-2021}. To gather the required knowledge, professionals may need to consult a variety of sources, such as clinical records, research papers, and medical guidelines in healthcare, or legal codes, court decisions, contracts, and investigative or forensic reports in law. 

Users' needs---such as searching for precedent cases in law or identifying relevant clinical trials in healthcare---have motivated the development of domain-specific \acf{ir} systems~\cite{moens2001innovative,breuker2004legal,hersh2003trec,roberts2016overview}. 
% These systems are designed to support complex queries and integrate multiple heterogeneous sources of information to aid expert decision-making~\cite{voorhees2014clinical,liu2023domain}.

In this thesis, we consider the legal domain as representative of knowledge-intensive domains, especially focusing on civil court judgments---i.e., the documents which conclude a civil trial and contains the parties, the facts, and the motivated final decision---and investigative chat logs---i.e., the chat messages extracted from seized suspects' phones.

However, the legal domain is characterized by specific challenges for \ac{ir}. According to the law, the theft of similar products may be regulated with different law articles, according to the price of the stolen item~\cite{feng-etal-2024-legal}. Consequently, two cases that appear similar at first glance may be, instead, quite different in the eyes of the law. The concept of relevance also heavily depends on the outcome of the trial~\cite{SANSONE2022101967}.
Furthermore, this domain is characterized by a peculiar language that follows implicit structure, which however can vary depending, for instance, on the judge that is writing the document~\cite{SANSONE2022101967}. Documents are often long in size and heterogeneous in nature~\cite{SANSONE2022101967}---for example, the data acquired during an investigation may include legal contracts, chat messages, posts on social media, or even pictures, and audio recordings.

Another important challenge for using \ac{ir} systems, or any other automated system in this domain, is the compliance with the regulations, such as respecting principles like transparency, non-discrimination, and privacy~\cite{SANSONE2022101967}\todo{magari cita direttamente regulations}.
\todo{time challenge, see feng. section aware,... aggiungi magari in legal o in soa. o quando definisco le proprietà}

\todo{say something about legal case retrieval systems}
\todo{speak of more use cases and why case ir is not enough}
% mention graphrag and other proposals (also not rag) that create an intermediary representation of the information
% idea di ordinamento (table of contents gerarchica, ir, ...)
% integrazione conoscenza
While, specialized methods for legal case retrieval do exist, the focus of this work is to achieve a level of generalization that allows the system to work in different scenarios, also including heterogeneus investigative data. Indeed, in this case there are no solution to achieve this????\todo{verify, cite, also from claw}
% siamo più generali di legal case search
\todo{proprietà da rispettare. una può essere anche la generalizzabilità. che poi riprende adattabilità in claw}

My proposal is to use entities, which are of undeniable importance in knowledge-intensive domains, think at law articles for the legal, or diseases and sympthoms for the healthcare, to organize the information from the data, for the final objective of helping final users in fulfilling their information need---also proposing novel information access applications.

% then say my proposal is similar 


% esempi business cases sul legal dove ai puo aiutare
% search engine / search bar

% ai in legal / kg intensive . integration tramite entità
% cosa si fa e challenges
% proprietà della soluzione

% immagine

\subsection{Context}

In Knowledge-intensive domains, such as law and criminal investigations, noticeable attention is given to the people, locations, or organizations mentioned in judgements, or investigative data.
\todo{entity/named entity definition}
Judges, layers, or investigators, indeed, often seek information around particular entities, but the vastness, variety, and ambiguity of the examined documents represent a challenge for a prompt and complete understanding~\cite{real_batini2021semantic,pozzi_wise_2025,BELLANDI2024,aixia_2023}, often resulting in time-consuming manual work.\todo{citare altro?}\todo{leggi il paper batini. find additional citations}
\todo{there is interest in using technology to help extract, organize and access information. we can see this by looking at the recent interesting in applying AI/ML techniques to...}
Recent interest, indeed, has been driven towards extracting and organizing knowledge for supporting the exploration and interpretation activities by the end users~\cite{kejriwal2018,szekely2015building,kejriwal2017investigative}.\todo{aggiungere altro: sono tutti dello stesso gruppo}

\subsection{Extraction}

Extracting entity-based knowledge from textual documents has been the focus of several studies, dating back to \todo{}. Researchers have been focusing on different aspects of the extraction process, from the identification of which words refer to a named entity---named entity recognition (NER)---to the curation of knowledge bases to maintain and serve information about knwon entities~\cite{}. And then the connection of the identified mentions of entities to the correct record in the knowledge base---named entity linking (NEL)~\cite{}.
\todo{in this work we relax the definition of named entity to entity in general}  

\subsubsection{Specific language}

However, the specific language used in these domains additionally hardens the task of extracting knowledge from domain-specific data~\cite{real_batini2021semantic,pozzi_wise_2025,BELLANDI2024,aixia_2023}.\todo{miei paper. ma sarebbe meglio aggiungere altre citazioni. forse anche kejriwal e sevgili possono essere citati}
In a criminal investigation, analyzing chat data is often necessary, for instance, to reconstruct people whereabouts, but this data usually contains typographic errors, slang, aliases, or abbreviations~\cite{}. This often leads to poor extraction performance when using \say{general domain} models trained on a different data distribution, often web-based data~\cite{}.
\todo{similar conclusions for legal judgements}
\todo{find evidence that this is a challenge from the literature. someone who had low performance.}

\subsubsection{Incremental}

Futhermore---while benchmark data often refer to publicly known entities---investigative data and legal judgements often mention people who are not described in any public knowledge base~\cite{}, such as Wikipedia~\cite{} or Wikidata~\cite{}. \todo{attento a wikidata vs wikipedia, KB vs KG}
As a consequence, these documents require to detect which mentions refer to unknown entities and to represent and add the new entities, incrementally, to the knowledge base for using them when linking future documents.
Recent interest has been driven towards addressing this problem~\cite{Heist2023nastylinker,kassner-etal-2022-edin}
\todo{cita lavory su NIL: nasty edin, iowa}
\todo{also here show why this is a challenge and what was/is missing in the literature}

\todo{mancanza di dataset}

\todo{summarize the challenges? nil pred, clustering? non ci ho lavorato molto}

\subsection{Access}
Alongside extracting knowledge from domain-specific documents, users require frameworks to access and explore the extracted knowledge to fulfill their information needs.

\subsubsection{UIs}
Recent work demonstrate the imporance of user interfaces for this objective~\cite{}, for instance\dots
\todo{citare kejriwal, sevgili e altri}
\subsubsection{QA}




\section{Objectives and Motivation}
\todo{prima di qui parlare bene dei problemi: cosa manca(va)????}
The main objective is fulfilling the information need of users in knowledge-intensive domains.
\todo{assicurati di disctere che questa cosa non era già possibile e di mostrarne i problemi che esistevano}
The main objective can be divided in different sub-objectives that correspond to the contributions of my Ph.D. research.
\begin{itemize}
    \item create a dataset and a baseline for incremental EE
    \item analyze existing EE algorithms on domain-pecific domains, evaluating different adaptation techniques
    \item build prototypes to allow users to explore domain-specific documents with the extracted knowledge
    \item build an efficient QA system for asking question on external knowledge in low-resource scenarios.
\end{itemize}
\todo{probabilmente si possono espandere}
\todo{esiste una piattaforma come dave? cosa esiste che va in conflitto con i miei contributi?}


\textbf{Motivations}. Properties of the focused domains. privacy is more a reason for some properties like the local execution of models.
\begin{itemize}
    \item privacy: not always possible to use APIs --- local
    \item scalability: to process large number of documents --- scalability, efficiency, batch-processing
    \item specificity of the domains --- adaptability??? I did not work on this... except aixia paper
    \item incrementality
\end{itemize}