%\pagestyle{mystyle}
\chapter{Knowledge Extraction in the Legal Domain}
\todo{kg because in wise it is not only entities but also relations from chat metadata}
\todo{add dave}
% \section{User interfaces}
% \subsection{Document Explorer}
% \subsection{Knowledge Editing to Correct Machine Errors}
% Human-in-the-Loop
% \subsection{Faceted Search}
% \subsection{Question Answering with Retrieval Augmented Generation}

% \section{Question Answering with ReFactX}

As anticipated in \Cref{sec:background:ai4legal}, \ac{ai} systems can be applied to the legal domain for addressing the analysis of vast amount of documents, that can come from heterogeneous sources.
During my Ph.D. I worked with two types of documents from the Italian legal domain: civil judgements and investigative chat logs. In both cases, the main problem addressed is the difficulty that the users---such as judges or prosecutors---face in analyzing the high number of records. Indeed, my co-authored contributions seek to empower final users with tools that allow them to efficiently explore and retrieve relevant documents or chat messages.
These proposals have been published in Computer Law \& Security Review (Vol. 52)~\cite{BELLANDI2024}, in the proceedings of the 22nd International Conference of the Italian Association for Artificial Intelligence (AIxIA 2023 -- Advances in Artificial Intelligence)~\cite{aixia_2023}, and of the 25th International Conference on Web Information Systems Engineering (WISE 2024)~\cite{pozzi_wise_2025}.

Data in this domain, as described in \Cref{sec:background:ai4legal}, often contain personal data, and privacy regulations imply important challenges such as the difficulty in obtaining labeled data for evaluating or training \ac{ee} systems. Furthermore, the language used in documents---such as civil judgments---differs from daily language~\cite{ccetindaug2023name_legal_turkish}, while open datasets used for training neural models, such as Common Crawl~\cite{commoncrawl} or The Pile~\cite{gao2020pile800gbdatasetdiverse}, are generally obtained from Web data. Another consequent issue is the impossibility of using APIs, such as OpenAI ones\footnote{https://openai.com/api/}\todo{mention apis, openai,... in llms}, with legal data that contain personal information---unless previously anonymized or pseudonymized. Moreover, using these APIs may involve transferring sensitive data outside the European Union, raising additional compliance concerns~\cite[Article~44]{gdpr2016}. As a result, institutions are often forced to rely on local models, which, while enabling full control over data storage and security in line with \textcite[Article~32]{gdpr2016}, can be limited by the local infrastructure.

In the remaining of this chapter, I present the contributions of my Ph.D. work on entity extraction in the Italian legal domain.
\todo{summary introduttivo}
With respect to civil judgments, due to the lack of labeled domain data, a benchmark dataset has been annotated with a semi-automatic approach and a \ac{nel} pipeline, inspired from the one introduced in \Cref{sec:incrementalel:pipeline}, has been evaluated on the labeled data~\cite{BELLANDI2024}. Additionally, different adaptations techniques have been studied to improve \ac{ner} performance on legal data~\cite{aixia_2023}, discussing both the accuracy and the required computation.

For criminal investigation data, we focused the extraction and organization of the entities from instant messaging data. Messages---including voice transcriptions---have been modeled in a knowledge graph that capture the network of contacts of the main suspect. Furthermore, an improved version of the \ac{nel} pipeline from \Cref{sec:incrementalel:pipeline} have been evaluated on data---previously labeled with a similar procedure to the one applied to legal judgments~\cite{pozzi_wise_2025}.

These works\todo{works plurale?} also include contributions in \acl{ia}, which are build upon the knowledge extracted from the domain data, and are described later in the next \Cref{chap:legalia}.

In summary, the main contributions in legal \ac{ee} of this Ph.D. are the following.
\begin{itemize}
    \item Creation of annotated benchmark datasets for civil judgments and investigative chat data, enabling structured evaluation of entity extraction methods.
    \item Development and evaluation of \ac{ner} and incremental \ac{nel} pipelines tailored for legal and investigative domains, including a study on domain adaptation techniques.
    \item Knowledge organization using knowledge graphs to represent entities and relationships from investigative chat records.
\end{itemize}

\todo{prima rw legal ie and italian}