\section{Data Integration Architectures for the Legal Domain}
\label{sec:sota:arch}
% \section{State of the Art on Data Integration and Management Architectures for Knowledge-Intensive Domain}

% some kms for investigations refer to sharing knowledge across individuals. e.g. tell others how i did this
% kms is not relevant!
% data management system

% \todo{maybe cite some requirements for legal: eg.g. semi automatic, traceability}
% % knowledge management
% \Acf{km} refers to the processes of generating, organizing, disseminating, and re-using an organization's information and expertise. A \emph{\acf{kms}} is the technological infrastructure that supports these processes by acquiring knowledge from internal or external sources, structuring it in a persistent and queryable form, and enabling efficient retrieval, sharing, and controlled evolution over time~\cite{computers12040072}.


% they are not \acp{kms} but do something related \eg legal case retrieval,...


% \paragraph{ai/nlp for legal soa. judgements and similar}

% Several systems for assisting legal professionals have been developed over time~\cite{amato2008,buey2016,breit_legal_permits_2020,HAN2024121052legalasst,human-ai-system-legal-precedent-search-china}.
% Early work by \textcite{amato2008} introduced a document management system for the legal domain that leverages ontologies and entity extraction, focusing primarily on notary documents.
% \textcite{buey2016} focused on a wider range of legal documents---such as public acts and private agreements---by employing ontologies that model document structures and specify which entities should be extracted from each section.
% \textcite{breit_legal_permits_2020} later explored the automation of processes related to legal permits\todo{add description once verified}.
% More recent work include \textcite{HAN2024121052legalasst,human-ai-system-legal-precedent-search-china}. The former proposed \emph{LegalAsst}, a system designed to enhance court productivity by integrating \acp{llm}, entity recognition, graph-based visualizations of trial facts, and decision-tree representations of judicial reasoning.
% The latter developed a prototype for legal case retrieval that guides professionals interactively through a faceted search interface---an exploratory search paradigm that organizes results along multiple dimensions, or facets~\cite{ARMENTANO2014_nlp_faceted,hearst_2006_faceted}, such as case type, court, date, or involved parties---recommending alternative search terms to support precedent discovery.

% While these work develop ontologies or multiple services including extraction and visualization tools, they do not focus much the architecture design or implementation.


%%%%%%%%%%%%%%%%%%%%555

% Several architectures have been proposed for integrating and managing information from legal documents. While they focus on different document types, and they address different problems they share a common motivation, that is the fact that the legal profession often need to perform time-consuming manual operations on vast document collections.

% In this section we review and categorize relevant approaches according to different dimensions corresponding to the requirements deriving from the legal domain, which have been introduced in \Cref{sec:intro}.

% With respect to generalizability, not all the approaches consider heterogeneous documents, but still recognize the need to generalize to different styles from different judges.

% In terms of traceability, some systems are developed to be traceable. With some also proposing interactive or step-by-step assistance over full automation.

% Finally, scalability....


Several architectures have been proposed to integrate and manage information from legal documents, motivated by the fact that legal professionals must process large collections of heterogeneous material under significant time constraints~\cite{bellandi2025streamlining,breit_legal_permits_2020,beniamino_strada_2022,HAN2024121052legalasst,human-ai-system-legal-precedent-search-china}.
% Among these, three lines of work are most closely related to the present thesis. \textcite{breit_legal_permits_2020} articulate architectural requirements for legal data processing (provenance, HITL correction, knowledge integration). \textcite{P_rez_2021} design a microservice-based pipeline for heterogeneous investigative evidence. \textcite{bellandi2025streamlining} present a platform for court decisions based on entity extraction, graph storage, and API-exposed pipelines for analytics. These will serve as the closest reference points.

To assess how existing approaches relate to the challenges of the legal domain (\Cref{sec:intro}), we organize the review along five dimensions: (i) generalizability to heterogeneous legal sources; (ii) traceability and verifiability of derived knowledge; (iii) error-correction and human oversight; (iv) scalability; and (v) support for downstream information-access tasks. Finally, we introduce a dedicated subsection on architectural choices, where we summarize the main design strategies adopted in the literature (e.g.\ microservices, service-oriented integration, provenance logging, orchestrators), independently of the tasks they support.
% most similar work

Among the reviewed approaches, three lines of work are most closely related to the present thesis. \textcite{breit_legal_permits_2020} do not implement a system but articulate architectural requirements for legal data processing, including traceability, provenance, human oversight, and integration of external knowledge. \textcite{P_rez_2021} address heterogeneous investigative evidence through a modular microservice pipeline that integrates multiple extractors. \textcite{bellandi2025streamlining} propose a platform for court documents that combines entity extraction, a graph-based registry, and API-exposed pipelines for analytics and exploration. These works are therefore the nearest reference points; the remainder of this section places them, together with other proposals, within the dimensions introduced above.

\subsection{Generalizability}
% Architectures for legal data management differ in the scope of documents they support. Early work focused on specific sub-domains: \textcite{amato2008} targeted notary documents, while \textcite{buey2016} extended the same architectural rationale to public acts and private agreements by modelling their internal structure and extracting section-specific entities. Other systems concentrate on legislation, such as Eunomos~\cite{boella2016eunomos}, which acquires statutory texts, preserves temporal versions, and links legal concepts to legislative fragments. These works show that even within a single category of legal sources, variation in drafting style and structure must be handled.

Different approaches have addressed different document types separately. \textcite{amato2008} focus on notarial documents, while \textcite{buey2016} extend a similar strategy to public acts and private agreements by modelling their internal structure and specifying which entities to extract from which section. \textcite{breit_legal_permits_2020} make the generalization requirement explicit, noting that heterogeneity arises even within a single document class because texts are produced by different legal entities with different drafting practices.


A different line of work considers judicial documents and investigative sources. \textcite{bellandi2025streamlining} address court decisions and procedural documents, integrating extraction services and analytics. Closer to investigative workflows, \textcite{szekely2015building,kejriwal2017searchengine,kejriwal2018} and \textcite{P_rez_2021} propose architectures for heterogeneous evidence, including pages from the dark-web and multimedia, organized via ontologies or microservice pipelines. In summary, generalizability is a recognized requirement, though most systems target one class of sources rather than a unified handling of heterogeneous legal and investigative documents.

\subsection{Traceability and Verifiability}

Several systems support tracing structured information back to its textual origin, although this is addressed with different depth. \textcite{breit_legal_permits_2020} introduce a provenance manager that records data transformations for auditing, and systems in the investigative domain such as \textcite{szekely2015building,kejriwal2017searchengine} record the source document for every node and edge in the constructed knowledge graph. Other systems, although not formulated around traceability as a requirement, still implement mechanisms that implicitly support it: in Eunomos~\cite{boella2016eunomos,boella2019semi}, concepts and roles are explicitly linked to legislative fragments, and \textcite{bellandi2025streamlining} adopt an annotation model that links extracted entities to document spans. \textcite{HAN2024121052legalasst} likewise incorporate a human-intervention interface allowing users to inspect, verify, and adjust automatically derived analyses.

In precedent exploration, \textcite{human-ai-system-legal-precedent-search-china} favor step-wise interaction over opaque automation, making intermediate steps inspectable. Overall, explicit traceability is directly addressed in a subset of systems, especially those motivated by auditing or investigative constraints, while others achieve only partial or implicit forms of traceability within the scope of their workflow.



\subsection{Error Correction and Human-in-the-Loop}

Human oversight is supported in several systems. Eunomos~\cite{boella2016eunomos} uses semi-automatic population, with users validating suggested annotations; \textcite{boella2019semi} extend this with \ac{hitl} correction for classification and extraction. \textcite{breit_legal_permits_2020} provide a correction interface for legal permits.

More recent systems couple extraction with interactive control. LegalAsst~\cite{HAN2024121052legalasst} includes a human-intervention module that exposes automatic prediction for review and adjustment. \textcite{human-ai-system-legal-precedent-search-china} explicitly present their system as AI assistance rather than automation, employing step-by-step assistance to ensure that decisions remain under human control.




\subsection{Scalability}

Several proposals address scalability and robustness through architectural choices. 
A common strategy is the adoption of microservices~\cite{P_rez_2021,buey2016,MorenoSchneider2021LynxCuration,RodriguezDoncel2019LynxDMP,bellandi2025streamlining,breit_legal_permits_2020}, which support distributed execution, modular replacement of components, and incremental extensibility. 
Lynx~\cite{MorenoSchneider2021LynxCuration,RodriguezDoncel2019LynxDMP} additionally employs an orchestrator and a validation controller to coordinate microservices.

Message-based coordination appears in \textcite{bellandi2025streamlining,P_rez_2021}, where pipelines are orchestrated via the publish-subscribe paradigm~\cite{pubsub_2003}, enabling asynchronous scaling of components.
Finally, \textcite{szekely2015building,kejriwal2017searchengine,kejriwal2018} explicitly consider scalability in the implementation choices for their algorithms and services, such the search engine.



\subsection{Support for Downstream Information Access}

Downstream \acl{ia} use cases are supported to different extents. \textcite{human-ai-system-legal-precedent-search-china} provide \acp{ui} for faceted search and interactive retrieval for precedents; \textcite{szekely2015building,kejriwal2017searchengine,kejriwal2018} expose faceted and map-based visualizations for investigative data; Eunomos~\cite{boella2016eunomos} supports semantic navigation in legislation.

LegalAsst~\cite{HAN2024121052legalasst} provides graph-based visualizations and decision summaries. \textcite{bellandi2025streamlining} support semantic document visualization, enriched with entities and concepts, as well as an \ac{ui} for analytics. Investigative systems~\cite{P_rez_2021,pirzada2024framework} offer dashboards and communication graphs without generalized retrieval or QA.


% \subsection{Architectural Choices and Synthesis}

\subsubsection{}
Beyond their specific functional goals, the reviewed systems differ in how they operationalize integration and control. 
Many adopt modular, service-oriented architectures~\cite{buey2016,P_rez_2021,MorenoSchneider2021LynxCuration,RodriguezDoncel2019LynxDMP,bellandi2025streamlining,breit_legal_permits_2020}, which facilitate distributed execution, independent service updates, and incremental extensibility. 
Lynx~\cite{MorenoSchneider2021LynxCuration,RodriguezDoncel2019LynxDMP,bellandi2025streamlining,P_rez_2021} complements this modularity with orchestration tools that coordinate the execution of services and maintain workflow consistency. 
\textcite{breit_legal_permits_2020} explicitly introduce a provenance manager to record data transformations.

Several systems also specify how processed information is represented and stored. 
Knowledge graphs and graph databases are commonly adopted to represent extracted entities, relations, and legal concepts~\cite{amato2008,kejriwal2017searchengine,kejriwal2018,bellandi2025streamlining,boella2016eunomos,MorenoSchneider2021LynxCuration,breit_legal_permits_2020}. 
In parallel, some systems~\cite{bellandi2025streamlining,boella2016eunomos,P_rez_2021} describe dedicated storage modules for documents and annotations, relying on different back-end technologies such as relational databases~\cite{boella2016eunomos}, NoSQL stores~\cite{P_rez_2021,szekely2015building,kejriwal2017searchengine,kejriwal2018}, or combining both~\cite{bellandi2025streamlining}. 
Several architectures also integrate indexing or search-engine components~\cite{bellandi2025streamlining,szekely2015building,kejriwal2017searchengine,kejriwal2018,human-ai-system-legal-precedent-search-china} to support keyword-based retrieval as well as more advanced access patterns, such as faceted navigation or semantic search.

Extraction services usually provide \ac{ner}, \ac{nel}, and rule-based annotation modules that populate these backends~\cite{boella2019semi,P_rez_2021,bellandi2025streamlining,beniamino_strada_2022}, and their results are often exposed through web-based \acp{ui} for visualization or correction~\cite{boella2019semi,HAN2024121052legalasst,bellandi2025streamlining,beniamino_strada_2022}. 
This organization---comprising extraction, storage and integration, and exposure for information access---appears recurrent across several systems, even when their goals or document types differ.

\medskip
\noindent
Overall, these architectural patterns converge toward modular systems that can evolve over time, even though many are scoped to specific legal or investigative subdomains. 
Taken together, the reviewed approaches demonstrate progress toward individual requirements such as generalizability, traceability, or scalability, but no existing architecture satisfies all of them simultaneously, or to the extent required in this thesis.

Compared with \textcite{breit_legal_permits_2020}, the architecture presented in \Cref{chap:architecture} extends generalizability to a broader range of heterogeneous documents, including those from investigative contexts. 
It also supports consolidated knowledge that goes beyond mention-level recognition (\acl{ner}) to entity-level consolidation through \acl{nel}, NIL prediction, and clustering, thereby enabling entities to serve as more powerful and semantically meaningful interlinks across diverse document collections.

Similarly, with respect to \textcite{bellandi2025streamlining}, our proposal pursues broader generalization and places stronger emphasis on error correction and traceability. 
Finally, relative to \textcite{P_rez_2021}, our architecture emphasizes traceability, verifiability, and error correction and supports entity-level knowledge consolidation, beyond mention-level, for integrating data.





%Also we keep a minimum necessary design of the architecture and not consider orchestration or extraction tools as part of it.

This observation motivates the design introduced in \Cref{chap:architecture}, which integrates these principles---including traceability, verifiability, and human oversight---into a unified, entity-centric architecture that supports heterogeneous data from both legal and investigative contexts, while remaining compatible with \acp{ui} for advanced \acl{ia} use cases, such as the ones defined in \Cref{sec:intro}.





% \todo{continua qui}
% % These choices indicate a tendency toward modular and auditable integration rather than monolithic processing, but they are usually applied within specific legal sub-domains and without a unified representation layer supporting both extraction and downstream information access over heterogeneous legal and investigative sources.

% \subsubsection*{Synthesis}

% Existing work addresses the five dimensions individually but not jointly. Generalizability is handled within domain-specific systems; traceability and \acl{hitl} mechanisms exist but are not propagated across heterogeneous sources; scalability is obtained in specific pipelines without unifying representation; downstream information access is supported in isolated settings rather than on the same integrated layer. These observations motivate the architecture proposed in \Cref{chap:architecture}, which is designed to satisfy the same dimensions simultaneously and to serve as the substrate for retrieval, navigation, analytics, and grounded question answering over heterogeneous legal and investigative data.





% %%%%%%%%%%%%5
% \paragraph{old}
% Several systems have been proposed to assist legal professionals in accessing, organizing, or interpreting legal material. %~\cite{amato2008,buey2016,breit_legal_permits_2020,HAN2024121052legalasst,human-ai-system-legal-precedent-search-china}.
% Early work by \textcite{amato2008} introduced a document management system for notary documents combining ontologies with entity extraction and regular expressions to extract triples and store them in an \ac{rdf} database. \textcite{buey2016} generalized this idea to other legal documents---public acts and private agreements---by modeling their internal structure and specifying which entities to extract from each section. Also they propose a service-oriented architecture for integrating different extraction methods dynamically.


% \textcite{breit_legal_permits_2020} explored the partial automation of legal permit procedures and, rather than focusing on a concrete implementation, articulated a set of architectural requirements for such systems. They argue that generalizability is essential due to the heterogeneity of input data produced by different legal entities, and that integration of external knowledge sources---such as legal norms or case law---is necessary to contextualize and enrich permit data. The design also requires \acl{hitl} error correction through a dedicated \ac{ui}, enabling users to inspect and revise system suggestions, with the expectation that these corrections could be leveraged to reduce future errors. In addition, they include auditability as a requirement for monitoring the system after deployment: to support this, they propose to use a dedicated provenance manager that logs every system operation on data---\eg data transformations into triples---thereby aligning with the traceability requirements discussed in \Cref{sec:intro}. In summary, their architectural proposal is composed of extraction services, an error-correction \ac{ui}, a \acl{kg}, and a provenance manager for auditing. 


% % \paragraph{Proper legal data management systems.}


% % \paragraph{proper legal data management systems}
% \todo{organizza per mostrare challenge, obiettivi e casi d'uso}

% Eunomos~\cite{boella2016eunomos} is a legal data management system that integrates legislative XML---used to represent statutes in a structured and versioned form---with legal ontologies encoding legal concepts and relations (\eg \emph{replaced-by} for amendments). The platform automatically acquires legislation (\eg from Normattiva\footnote{\url{https://www.normattiva.it/}}), converts it to XML, extracts cross-references to build a network of norms, and maintains temporal versions of provisions. Concepts, prescriptions, roles, and sanctions are stored in an ontology and linked to specific legislative fragments, enabling semantic navigation and fine-grained queries. Population is semi-automatic: the system proposes annotations while still allowing manual correction. Targeted use cases include compliance in financial regulation and support for law firms and public bodies in tracking and structuring legal knowledge, as well as citizen-facing access to consolidated and concept-aware legislation. By coupling structured legislation with a concept layer, Eunomos goes beyond keyword-based retrieval and enables controlled, ontology-driven access to evolving legal norms.

% \textcite{boella2019semi} extend Eunomos with improved \ac{nlp} services to reduce the manual effort of knowledge population via semi-automatic classification of norms, concept and relation extraction, amendment extraction, and \acl{nel} using pattern-matching rules. A web-based \acl{hitl} interface is provided for validating the \ac{nel} outputs, and a similar \ac{ui} is envisioned for supporting semi-automatic refinement of amendment extraction.

% % error correction? traceability?
% \textcite{beniamino_strada_2022} acknowledge the importance of organizing the documentation attached to legal trials and propose a workflow for automatically determining the proposability of the instances presented to Courts, with the goal of reducing the burden on judges. They build a combined ontology from proposability, juridical, and road-accident ontologies, and train an ad-hoc \ac{ner} model to recognize domain-specific entities in trial documents (\eg injured party, tax identification code, description of the facts, lawyer, statement). Proposability is then computed by logical inference over the extracted entities. This work shares similarities with some of the contributions of this \phd, since it adopts an entity-centric approach and employs \ac{ner} for knowledge extraction, as in \textcite{BELLANDI2024,aixia_2023,pozzi_wise_2025}. It also involves the annotation of a dataset for training \acl{ner}, as in \textcite{BELLANDI2024,pozzi_wise_2025}. In addition, the authors provide user-facing \acp{ui} for visualizing proposability analysis and for displaying \ac{ner} annotations highlighted on documents side-by-side with the ontology, similarly to the \acp{ui} presented in \textcite{BELLANDI2024,pozzi_wise_2025,Agazzi2025DAVE}. However, the work does not describe a general legal architecture; it rather specifies a workflow for extracting ontology concepts with \ac{ner} to support the specific proposability task.

% \textcite{balansay2024development} focus on notarial documents and aim to reduce the time-consuming manual operations carried out by professionals for document management. They develop a management system with a database for storing documents, organizing them into folders, searching them, and sharing them with colleagues, providing all these services through a web-based \ac{ui}.

% \textcite{enase22,icaart23,pham2025improving} focus on Vietnamese legislation and introduce the Legal-Onto model~\cite{enase22}, an ontology framework structured around concepts, relations, and inference rules. A conceptual graph links key phrases to legal concepts, supporting semantic representation and retrieval. In later work~\cite{pham2025improving}, the authors address the problem of keeping the ontology up-to-date by parsing and structuring legal documents, extracting key phrases, and reconstructing the conceptual graph, as well as comparing different versions of laws through vector-based similarity. The overall architecture consists of the legal database---where documents are parsed and segmented into sections, articles, clauses, etc.---and the conceptual graph built from them, along with user interfaces for search and exploration~\cite{icaart23,enase22}.


% Lynx~\cite{MorenoSchneider2021LynxCuration,RodriguezDoncel2019LynxDMP} provides a microservice-based architecture for building and exploiting a multilingual Legal Knowledge Graph for compliance. Documents from heterogeneous legal sources are assigned URIs, enriched with metadata, decomposed into parts, and exposed through APIs. The platform offers services for search (including over annotations), question answering, terminology access, and workflow execution over documents via an orchestrator that coordinates document-level operations, in a spirit similar to the automation pipelines in \cite{BELLANDI2024}. A validation controller ensures integrity by enforcing that stored documents conform to the data model (text, metadata, identifier, type). The Legal Knowledge Graph aggregates and interlinks legal and regulatory data across jurisdictions and languages, leveraging Linked Data principles. Although the broader legal graph on the \Ac{web} is unbounded, Lynx materializes and serves the core entities relevant for its services, primarily legal documents (acts, directives, standards), concept schemes, and selected organizational entities. A document manager is designed and implemented for storing the Legal Knowledge Graph and the processed documents. 


% % heterogeneus
% On the other hand, approaches supporting criminal investigations have targeted illicit web data~\cite{kejriwal2017searchengine,kejriwal2018}, particularly in efforts to combat human trafficking. These works propose an entity-centric scalable architecture that constructs a \acl{kg}---according to a domain-specific ontology---and provides advanced \acl{ia} \acp{ui}, such as faceted search interfaces and map-based visualizations. Although the architecture could be generalized to other types of documents---with a different ontology---they did no experiment with multimedia data and implement the extraction services for specific challenges of illicit data, such as information that is often hidden in plain sight, such as phone numbers deliberately obfuscated by mixing digits and words.

% % both focus also on scalability
% \textcite{P_rez_2021} proposed a modular platform for analyzing vast and heterogeneous data---including multimedia content---from multiple sources. Their system is built upon an ontology composed by the entities \texttt{Person}, \texttt{Location}, \texttt{Organization}, \texttt{Equipment}, and \texttt{Event}, and adopts a scalable microservices-based architecture~\cite{microservices_2016} that integrates components such as speech-to-text and keyword extraction modules to populate a \ac{kb}. While the architecture is both modular and generalizable, their experiments rely primarily on keyword-based extraction techniques, which provide limited semantic understanding and require manual intervention.


% \textcite{bellandi2025streamlining} present an evolution of the work in~\textcite{BELLANDI2024}, which constitutes one of the contributions of this \phd The resulting platform enforces streamlined management and analysis of legal documents (\eg court decisions, orders) to serve different legal actors through service pipelines. Its architecture is modular and aligns with some principles underlying \ref{ob:architecture}. The architecture is composed of text-processing services---including \ac{ner}, \ac{nel}, and concept extraction---storage services for documents, entities, and annotations. A graph database is used as an entity registry, exposed through dedicated APIs. Pipelines are orchestrated through a Kafka-based mechanism and expose all services through standard APIs. The platform supports entity-based analytics (\eg entity distributions, paths between entities across documents) and concept-based document exploration, enabling navigation, search, and statistics over the corpus using extracted entities and concepts. Front-end modules provide \acp{ui} for exploration and analytics, extending the interfaces in~\textcite{BELLANDI2024}.

% However, the work focuses more on the operationalizing, rather than to improve an architectural model. In contrast, the proposal in \Cref{chap:architecture} emphasizes generalizability to heterogeneous documents, including investigation chat logs. Also, \textcite{bellandi2025streamlining} does not address error correction (which is relegated to future work) and does not explicitly cover traceability or verifiability, although it adopts the document--annotation--entity model and suggests human-in-the-loop supervision.

% Other recent systems incorporate \ac{nlp} and \acp{llm}. \emph{LegalAsst}~\cite{HAN2024121052legalasst} integrates entity recognition, anonymization, and a dedicated visualization \ac{ui} that displays graph-based representations of trial facts and decision-tree summaries of judicial reasoning to assist judges. The system is designed to be extensible, traceable, and controllable via interactive human correction, while retaining explainability through the explicit visualization of the inferred decision path.


% \textcite{human-ai-system-legal-precedent-search-china} focus on case retrieval and identify three design requirements for human-AI collaboration in precedent search: visualized faceted search---\ac{ai} assists users by suggesting or grouping related search terms, but the user controls the facets and thresholds; semantic zoom---users can move across levels of detail in a precedent (facts, reasoning, citations) with optional support such as summarization or citation tracing, and can bookmark items to bridge exploration and documentation; and step-by-step assistance---the system provides procedural help at a low level, ensuring that legal conclusions remain under human control. Overall, \ac{ai} is explicitly positioned as an assistive layer that enhances the search and reading process without taking over legal judgement.

% \todo{summarize interesting aspects of these works wrt my contrib}



% \todo{bellandi streamlining}~\cite{bellandi2025streamlining} this paper is an evolution of \textcite{BELLANDI2024} which is one of the contributions of my \phd Our collogues continued the work improving it essentially. this is why this work shares similarities with the architecture proposed for \ref{ob:architecture}.

% designed to enforce the streamlined
% management of legal documents, such as court decisions or
% orders, to serve a multiplicity of legal actors and processes.
% The IDJ platform consists of a set of modules, repositories,
% and data flows that interoperate to realize service pipe-
% lines enforcing legal document analytics and exploration

% The IDJ platform provides a text processing mod-
% ule, implementing basic services for document processing
% typically used at ingestion time (\eg text/metadata storage,
% tokenization, full-text search), and a NLP module, imple-
% menting advanced natural language processing services,
% used for knowledge extraction from ingested documents
% (\eg named entity extraction, named entity linking, concept
% extraction). End-user analytics and exploration processes
% over legal documents are realized by service pipelines that
% accomplish the process goal through the interoperation of
% selected text processing/NLP modules, repositories, and data
% flows of the IDJ platform.

% the entity-based document analytics, which enforces
% entity-based statistics generation over the document reposi-
% tory, thus allowing for example to analyze entity distribution
% over the document corpus; (ii) the concept-based document
% exploration, which enforces the document repository explo-
% ration according to concept-based similarity criteria. The
% IDJ platform is flexible, in that new service pipelines can be
% realized to enforce new stakeholders analysis requirements
% arising over time.

% graph database for the entity registry

% The storage layer exposes the EReg APIs to manage
% both the entity types (the EReg metamodel) and the entity
% instances

% Document texts and metadata
% are stored in an ElasticSearch instance, while annotations in
% a SQL database

% e back-end components and the em front-end compo-
% nents constitute the backbone of the proposed platform. In
% back-end components, we distinguish modules for text pro-
% cessing and NLP. About the text processing module, it can be
% considered a proxy for client utilities like indexing, embed-
% ding, and full-text search.

% Further services are included in
% the text processing module to process the incoming data at
% ingestion time, and to create manipulated versions of the
% original documents through activities like tokenization,
% lemmatization, cleaning, and filtering.

% Named Entity Recognition
% (NER) and Named Entity Linking (NEL), as well as concept
% extraction are also included

% %orchestration
% A Kafka queue is created when a NLP service is
% invoked, with all the information needed to get the proper
% data. All the NLP services must expose standard APIs for
% interaction with other platform components; they must read
% the Kafka queue and use the parameters found in it to obtain
% the input texts. In the end, the invoked NLP service passes
% back the output to the text processing module for storage in
% the annotation database and the entity registry.

% In front-end components, we extend our previous work in
% [2] and we distinguish modules for exploration and analyt-
% ics.

% service for brows-
% ing the corpus according to their common entities and/or
% concepts extracted by the NLP module. Analytics allows to
% examine the corpus through summary/statistical views built
% over data, such as for example the distribution of an entity
% or concept in the corpus, the shortest path (through docu-
% ments) between given concepts or entities, and the centrality
% of entities and concepts.

% reference data model charac-
% terized by documents, entities, and annotations

% document has text and metadata (key value)
% document has sections and chunks

% entities has type and features. information that allows to distinguish/identify an entity is called key and depende on the type (\eg fiscal code for person)

% graph-based concept exploration

% example concept: banking contract

% they use special entity types identified form the corpus of court judgment.

% concept-based document exploration


%%%%%%%%%%%%%%%%%%%%%


% \todo{paper architectural sota}

% \paragraph{other more specific legal systems}
% they are not \acp{kms} but do something related \eg legal case retrieval,...


% \paragraph{ai/nlp for legal soa. judgements and similar}
% Several systems for assisting legal professionals have been developed over time~\cite{amato2008,buey2016,breit_legal_permits_2020,HAN2024121052legalasst,human-ai-system-legal-precedent-search-china}.
% Early work by \textcite{amato2008} introduced a document management system for the legal domain that leverages ontologies and entity extraction, focusing primarily on notary documents.
% \textcite{buey2016} focused on a wider range of legal documents---such as public acts and private agreements---by employing ontologies that model document structures and specify which entities should be extracted from each section.
% \textcite{breit_legal_permits_2020} later explored the automation of processes related to legal permits\todo{add description once verified}.
% More recent work include \textcite{HAN2024121052legalasst,human-ai-system-legal-precedent-search-china}. The former proposed \emph{LegalAsst}, a system designed to enhance court productivity by integrating \acp{llm}, entity recognition, graph-based visualizations of trial facts, and decision-tree representations of judicial reasoning.
% The latter developed a prototype for legal case retrieval that guides professionals interactively through a faceted search interface---an exploratory search paradigm that organizes results along multiple dimensions, or facets~\cite{ARMENTANO2014_nlp_faceted,hearst_2006_faceted}, such as case type, court, date, or involved parties---recommending alternative search terms to support precedent discovery.


% \textcite{CHIEN2024105902} focus on Legal Judgment Prediction from indictments, \ie the task of predicting the applicable legal articles, charges, and sentences given the statement of facts. A criminal indictment typically contains the cause of charge, the narrative of facts and reasons, and the cited legal provisions and prosecutorial requests, providing a structured input for the prediction task.

% \todo{adele}
% \todo{non fanno nel ma alcuni organizzano in triple rdf. magari aggiungerlo}
% \todo{maybe order historically}
% \todo{add approaches for criminal investigations from claw}

% \paragraph{soa criminal}
% \todo{add whatsapp}



% \todo{paper architectural sota}

% \paragraph{other more specific legal systems}


% % \textcite{CHIEN2024105902} focus on Legal Judgment Prediction from indictments, \ie the task of predicting the applicable legal articles, charges, and sentences given the statement of facts. A criminal indictment typically contains the cause of charge, the narrative of facts and reasons, and the cited legal provisions and prosecutorial requests, providing a structured input for the prediction task.

% \todo{adele}
% \todo{non fanno nel ma alcuni organizzano in triple rdf. magari aggiungerlo}
% \todo{maybe order historically}
% \todo{add approaches for criminal investigations from claw}

% \paragraph{soa criminal}
% \todo{add whatsapp}
