%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{State of the Art on Question Answering with External Knowledge}
\label{sec:sota:qa}

\todo{add new work with constrained}

\begin{itemize}
    \item what is qa
    \item which qa we focus open? yes with big kg (wikidata)\cite{zaib2022conversationalqasurvey}
    \item external kg
    \item ke qa
\end{itemize}

Several approaches have been proposed to incorporate external knowledge into LLMs~\cite{gao2024retrievalaugmentedgenerationlargelanguage}. In our work, we focus on Question Answering (QA)---more specifically, on \textbf{Knowledge-Enhanced QA (KE-QA)}, where a model acquires external knowledge for answering questions.

Most of approaches can be classified as \textbf{input-based KE-QA}.
They rely on external tools, such as search engines or dense retrievers~\cite{karpukhin-etal-2020-dense}, which can be arranged in pipeline-based workflows~\cite{lewis2020retrievalaugmented,xu2023retrieval} or actively called by LLMs~\cite{yao2023reactsynergizingreasoningacting,NEURIPS2023_d842425e,lehmann2024beyond}. These approaches usually feed external knowledge as input (either in plain text or with dense embeddings) to the LLMs. Input-based KE-QA approaches, while effective in reducing hallucinations~\cite{shuster-etal-2021-retrieval-augmentation}, require additional models or services for retrieving external information that can introduce delays, especially when calling remote APIs, increase training complexity~\cite{sun2025zerosearchincentivizesearchcapability},
and raise substantially the number of input tokens~\cite{yang2024memory3}, leading to higher response time and resource usage.

Other approaches can be classified as \textbf{memory-based KE-QA}. They propose memory modules, separated from model parameters, which make the models' internal mechanisms interact with external knowledge, e.g., by using cross-attention to incorporate external vectors~\cite{yang2024memory3,wu-etal-2022-efficient}. While memory-based KE-QA approaches fuse external knowledge into LLMs' generation process more deeply, they typically require architectural modifications, complicating the use of pretrained LLMs.




Constrained generation~\cite{cao2021autoregressive,scholak-etal-2021-picard}, described in detail in Section~\ref{sec:constrained-gen}, is a promising technique to avoid the drawbacks of input-based and memory-based KE-QA approaches. It has been successfully applied to various tasks, including code generation~\cite{scholak-etal-2021-picard}, guaranteeing that the model produces syntactically correct code, as well as entity linking~\cite{cao2021autoregressive} and information retrieval~\cite{bevilacqua2022autoregressive}.
%
It has also been applied to instruction-following LLMs for several tasks that require the LLM to produce a specifically structured output, such as entity disambiguation~\cite{geng-etal-2023-grammar}.


Two recent studies~\cite{li2024decodinggraphsfaithfulsound,luo2024graphconstrainedreasoningfaithfulreasoning} proposed \textbf{constrained generation for KE-QA}, especially for guiding LLMs through paths from a knowledge graph (KG). However, they do not address the problem of accessing facts from large knowledge bases in a scalable way (800 million facts, in our case),  and still rely on entity linking systems for extracting question entities, similarly to input-based approaches. 
%
A first approach, \textit{Decoding on Graphs (DoG)}~\cite{li2024decodinggraphsfaithfulsound}, generates triples from a knowledge graph in a reasoning process similar to CoT~\cite{wei2022chainofthought} but grounded on KG knowledge, also allowing the model to alternate the constrained generation of triples and the reasoning with normal generation.
A second approach, \textit{Graph-Constrained Reasoning (GCR)}~\cite{luo2024graphconstrainedreasoningfaithfulreasoning} generates entire paths from a KG using a fine-tuned LLM to find answer hypotheses, and then asks a top-tier language model to give a final answer based on the paths.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\paragraph{KGQA}

%%%%% dog

Previous attempts on solving knowledge graph
question answering (KGQA) often involve seman-
tic parsing (Lan et al., 2022; Yu et al., 2023)
that transforms questions into logical queries (e.g.,
SPARQL) to be executed over KG for answers

(Lan and Jiang, 2020; Ye et al., 2022). These meth-
ods may require ground-truth logical queries for
training, and model-generated queries might be
non-executable due to syntax or semantic errors
(Yu et al., 2023; Luo et al., 2024; Mavromatis
and Karypis, 2024). Recently, researchers have
been exploring the joint use of KGs and large lan-
guage models (LLMs) for KGQA, leveraging the
advanced capabilities of LLMs for enhanced rea-
soning (Salnikov et al., 2023; Wu et al., 2023; He
et al., 2024). Various approaches (Jin et al., 2024;
Sun et al., 2024b) iteratively prompt LLMs to syn-
ergize with KGs to select relevant information step
by step (Ren et al., 2024). ToG (Sun et al., 2024a)
prompts LLMs to explore relevant facts hop-by-
hop using beam search on KG. Tree-of-Traversals
(Markowitz et al., 2024) is a zero-shot reasoning al-
gorithm motivated by Tree-of-Thoughts (Yao et al.,
2023). Instead of focusing on one hop at a time,
D O G uses the entire question graph as input, en-
abling reasoning over complete knowledge. Addi-
tionally, these prompting approaches overly rely
on the instruction-following abilities of LLMs, pos-
ing challenges when applied to compact LLMs.
Some studies (He et al., 2024) focus on training
specialized retrievers to extract key information
from KG and the retrieved knowledge as input to

LLMs. RoG (Luo et al., 2024) and GNN-RAG
(Mavromatis and Karypis, 2024) train in-domain
LLM and GNN retrievers respectively. StructLM
(Zhuang et al., 2024), on the other hand, trains gen-
eralist models over large amount of structured data
to augment the structured knowledge grounding
capabilities of LLMs. However, fine-tuning large
scale models is computationally expensive, and spe-
cialized training suffers from limited generalization
capabilities. Conversely, D O G employs a general
graph-aware decoding approach, eliminating the
need for external retrievers or specialized training.

%%%%%%%%%
% gcr
%reasoning
LLM reasoning. Many studies have been proposed to analyze and improve the reasoning ability
of LLMs (Wei et al., 2022; Wang et al., 2024; Yao et al., 2024). To elicit the reasoning ability
of LLMs, Chain-of-thought (CoT) reasoning (Wei et al., 2022) prompts the model to generate a
chain of reasoning steps in response to a question. Wang et al. (2024) propose a self-consistency
mechanism that generates multiple reasoning paths and selects the most consistent answer across
them. The tree-of-thought (Yao et al., 2024) structures reasoning as a branching process, exploring
multiple steps in a tree-like structure to find optimal solutions. Other studies focus on fine-tuning
LLMs on various reasoning tasks to improve reasoning abilities (Yu et al., 2022; Hoffman et al.,
2024). For instance, OpenAI (2024c) adopts reinforcement learning to train their most advanced
LLMs called “OpenAI o1” to perform complex reasoning, which produces a long internal chain of
thought before final answers.

%reasoning for qa
KG-enhanced LLM reasoning. To mitigate the knowledge gap and hallucination issues in LLM
reasoning, research incorporates KGs to enhance LLM reasoning (Pan et al., 2024). KD-CoT (Wang
et al., 2023) retrieve facts from an external knowledge graph to guide the CoT performed by LLMs.
RoG (Luo et al., 2024) proposes a planning-retrieval-reasoning framework that retrieves reasoning
paths from KGs to guide LLMs conducting faithful reasoning. To capture graph structure, GNN-
RAG (Mavromatis & Karypis, 2024) adopts a lightweight graph neural network to effectively re-
trieve from KGs. Instead of retrieving, StructGPT (Jiang et al., 2023) and ToG (Sun et al., 2024)
treat LLMs as agents to interact with KGs to find reasoning paths leading to the correct answers.

%rag
rag \cite{gao2024ragsurvey}

qa with cot \cite{yoran-etal-2023-answering}