\subsection{NIL Prediction}
\todo{maybe join with clustering}

\begin{definition}[NIL Prediction]
    Given an entity mention $m$ and a reference \ac{kr}, \emph{NIL prediction} is the subtask of \acl{nel} that identifies whether $m$ referes to an entity that does not exist in the \ac{kr}~\cite{entity_linking_with_kb_2015}. 
\end{definition}

According to some work \emph{NIL} stands for ``not in lexicon''~\cite{ILIEVSKI2020100565}. Another interpretation is that it simply refers to the English word ``nil'' that means ``nothing''~\cite{cambridge_nil}.

This task has also been called with different names: ``unlinkable mention prediction''~\cite{entity_linking_with_kb_2015}, ``out-of-knowledge-Base mention discovery''~\cite{dong_reveal_the_unknown_2023}, ``out-of-knowledge-graph entity detection''~\cite{moller2024entity}, ``unknwon entity discovery''~\cite{kassner-etal-2022-edin}, and ``emerging entity discovery''~\cite{hoffart_discovering_emergin_entities_2014}.

\textcite{zhou-etal-2024-gendecider} distinguishes between NIL prediction and ``none of the candidates'' detection, arguing that \emph{NIL} means the referred entity is not in the \ac{kr}, while ``none of the candidates'' means the referred entity has not been retrieved without assuming it is not present in the \ac{kr}. However, in practise, NIL prediction is often performed by predicting whether the top-ranked entity, returned by the \ac{nel} ranking step, is correct or not~\cite{entity_linking_with_kb_2015}. It is impractical and prohibitive to compare a mention with every entity in a \ac{kr} for ensuring the mentions is NIL. Thus, in this work, we consider the preactical assumption that if a mention does not refer to the top-ranked candidate(s), returned by the \ac{nel} system, the mention is NIL. Under this assumption there is no difference between NIL prediction and none-of-the-candidates prediction.

NIL prediction has frequently been overlooked in \acrshort{nel}
systems, as many approaches exclude NIL mentions from their evaluation datasets and
eventually leave this task to future works. In fact, among the 38 approaches\todo{verify the published version}
compared by the survey \citet{Sevgili2022NeuralEL} only 8 addressed NIL prediction.

% dico che le TAC hanno dato importanza al NIL
Early work on NIL prediction dates back to 2006, with \textcite{bunescu-pasca-2006-using} using a threshold-based approach.
Later, the Text Analysis Conference (TAC)\footnote{\url{https://tac.nist.gov/}} fostered the research in NIL prediction by introducing NIL prediction in its \acf{kbp} track starting from 2009 \cite{mcnamee2009overview}. Starting from 2011, participants were also required to cluster together those NIL mentions that referred to the same entity \cite{ji2011tac}. Organized by NIST\footnote{\url{https://www.nist.gov/}}, TAC aimed to advance research in \acrshort{nlp} and related areas by offering large-scale test collections and standardized evaluation procedures. Since 2008, it has been held annually, inspiring a variety of studies on NIL prediction.

Advancements in NIL prediction continued also in parallel to the TAC conference~\cite{gottipati-jiang-2011-linking,hoffart_discovering_emergin_entities_2014,hoffart_awakens_2016,wu2016exploring}, even if \textcite{gottipati-jiang-2011-linking} experimentation was run on the TAC-KBP 2010 dataset~\cite{ji2010overview}. 

\todo{moller paper reviewed a lot of approaches}
\todo{our contribution}


\todo{hoffart}
\todo{neel challenge}

\ac{nel} with NIL prediction can be framed as a classification problem with a reject option, which corresponds to NIL. According to \citet{Sevgili2022NeuralEL}, four main strategies are typically used:
\begin{enumerate}
\item Assigning NIL whenever candidate generation produces no possible entities.
\item Applying a threshold to the linking score, below which the mention is treated as unlinkable.
\item Introducing NIL as an additional candidate during the ranking stage, effectively creating a new class in the classification problem.
\item Training a separate binary classifier on mention-entity pairs, often with extra features such as the top linking score or \ac{ner} labels, to decide whether a mention is linkable or not.
\end{enumerate}

Some studies \cite{Zhang2012TheNE, Rao2013EntityLF} argue that methods \ref{nilm-threshold} and \ref{nilm-add-class} are essentially equivalent, although the latter avoids manual threshold tuning.

\todo{add recent studies}

In most cases, these techniques are incorporated into pipeline-based systems that treat NIL prediction and \acrlong{nel} as distinct steps~\cite{}. By contrast, other works have proposed joint models to reduce error propagation and leverage dependencies between the two tasks \cite{fahrni2013hits,martins-etal-2019-joint}.



\subsection{NIL Clustering}
\subsection{coreference resolution}
\todo{what did LLMs change?}
\subsection{Human-in-the-Loop}
